---
title: "hockeyR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hockeyR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting started

```{r setup}
library(hockeyR)
```

## `load_pbp()`

As mentioned on the [home](https://hockeyr.netlify.app/index.html) page, the main function of the `hockeyR` package is to load raw NHL play-by-play data without having to scrape it and clean it yourself. The `load_pbp()` function will do that for you. The `season` argument in `load_pbp()` is very accepting. You may use any of the following syntax when loading play-by-play data for the 2020-21 NHL season:

* Full season definitions (ie '2020-2021')
* Short seaosn definitions (ie '2020-21')
* Single season definitions (ie 2021)

To load more than one season, wrap your desired years in `c()`. That is, to get data for the last two years, one could enter `load_pbp(c(2020,2021))`.

## `get_game_ids()`

If you want to load play-by-play data for a game that isn't in the [data repository](https://github.com/danmorse314/hockeyR-data), or perhaps you just want a single game and don't need to load a full season, you'll first need to find the numeric game ID. The `get_game_ids()` function can find it for you as long as you supply it with the date of the game in `YYY-MM-DD` format. The function defaults to the current date as defined by your operating system.

```{r game_ids}
# get single day ids
get_game_ids(day = "2017-10-17")
```

You can instead supply a season to `get_game_ids()` to grab a full year's worth of IDs as well as final scores, home and road teams, and game dates for each game in the given season.

## `scrape_game()`

This function scrapes a single game with a supplied game ID, which can be retrieved with `get_game_ids()`. Live game scraping has yet to undergo testing.

```{r game}
scrape_game(game_id = 2020030175)
```

## `scrape_day()`

This is the backbone function that keeps the [hockeyR-data repository](https://github.com/danmorse314/hockeyR-data) up to date during the season. Supply a date (`YYY-MM-DD`) and it will scrape play-by-play data for all games on that day. Live game scraping is still awaiting testing.

```{r day}
scrape_day("2015-01-06")
```

If you can wait until the day after a game, the `load_pbp()` function is the only one you'll need. If you'd like to scrape the data yourself immediately following a game, the other functions discussed here will do the job for you.
