---
title: "Scraping hockey-reference.com"
author: "Daniel Morse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scraping hockey-reference.com}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/vignettes-",
  out.width = "75%",
  dpi = 750,
  fig.width = 7,
  fig.height = 4.67
)
```

```{r setup}
library(hockeyR)
`%>%` <- magrittr::`%>%`
```

###   Hockey-Reference scrapers

These functions scrape data from [hockey-reference.com](https://www.hockey-reference.com).

Grab every team's win-loss record in any season going back to 1918 with the `get_team_records()` function

```{r records}
get_team_records(1967) %>%
  dplyr::arrange(-w) %>%
  dplyr::select(team_name, team_abbr, season, overall, w, l, otl, st_points)
```

You can also get stats down to the player-level with `get_player_stats_hr()`. This function defaults to the player's career statistics, but you can enter a specific season or range of seasons as well. Note that the season references the year the specific season ended (ie the 2021-22 season should be entered as 2022)

```{r player stats}
get_player_stats_hr(player_name = "Wayne Gretzky", season = 1982) %>%
  dplyr::select(player, age, season_full, tm, gp, g, a, pts)
```

To get the basic counting stats for every skater in a given year, use the `get_skater_stats_hr` function. Like the other functions in this package, the `season` argument needs to be the end-year of the season you wish to scrape. 

```{r skater stats}
df2 <- get_skater_stats_hr(2022)

df2 %>%
  dplyr::arrange(-goals) %>%
  head()

```

You can use the data to make plots with actual team colors and logos as well using the `team_logos_colors` file included with the package.

```{r plot-example}
# add colors & logos
df3 <- df2 %>%
  dplyr::arrange(-points) %>%
  dplyr::slice(1:10) %>%dplyr::select(player, team_abbr, goals, assists, points) %>%
  dplyr::left_join(team_logos_colors, by = "team_abbr")

# make a bar chart
df3 %>%
  ggplot2::ggplot(ggplot2::aes(stats::reorder(player, -points), points)) +
  ggplot2::geom_col(fill = df3$team_color1, color = df3$team_color2) +
  ggimage::geom_image(
    ggplot2::aes(y = points + 2, image = team_logo_espn),
    size = .05, asp = 1.5
  ) +
  ggplot2::geom_text(ggplot2::aes(y = 2, label = player),
            color = "white", angle = 90, hjust = 0) +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::theme(
    panel.background = ggplot2::element_rect(fill = "black"),
    plot.background = ggplot2::element_rect(fill = "black"),
    panel.grid.major.x = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_text(color = "white"),
    title = ggplot2::element_text(color = "white")
  ) +
  ggplot2::labs(x = NULL, y = "Points",
       title = "Top-10 point scorers in the NHL in 2021-22",
       caption = "data pulled from hockey-reference.com using hockeyR")
```
Use the `get_rosters()` function to look up a team roster at season's end for any prior season. By default, it will only pull basic player info (name, age, height & weight, etc), but you can grab all the basic counting stats by setting `include_stats` to `TRUE`. Note as shown below that the `team` argument accepts both full team names or team abbreviations.

```{r rosters}
player_stats <- get_rosters(c("COL","Detroit red wings"), season = 2001, include_stats = TRUE) %>%
  dplyr::mutate(
    g_60 = 60 * g / toi,
    a_60 = 60 * a /toi,
    p_60 = 60 * pts / toi
  ) %>%
  dplyr::filter(toi >= 300) %>%
  dplyr::left_join(team_logos_colors, by = "team_abbr")

top_performers <- dplyr::filter(
      player_stats,
      p_60 >= dplyr::arrange(player_stats, -p_60) %>%
        dplyr::slice(10) %>%
        dplyr::pull(p_60)
      )

player_stats %>%
  ggplot2::ggplot(ggplot2::aes(a_60,g_60)) +
  ggplot2::geom_hline(yintercept = 60 * sum(player_stats$g) / sum(player_stats$toi),
             linetype = "dashed", color = "black") +
  ggplot2::geom_vline(xintercept = 60 * sum(player_stats$a) / sum(player_stats$toi),
             linetype = "dashed", color = "black") +
  #geom_point(aes(size = toi), show.legend = FALSE,
  #           color = player_stats$team_color_alt1, alpha = .8) +
  ggimage::geom_image(ggplot2::aes(image = team_logo_espn),
                      size = 0.07, asp = 1.5) +
  ggrepel::geom_text_repel(
    data = top_performers,
    ggplot2::aes(label = player),
    color = top_performers$team_color_alt1
  ) +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::theme(
    panel.background = ggplot2::element_rect(fill = "#708090"),
    plot.background = ggplot2::element_rect(fill = "#708090"),
    title = ggplot2::element_text(color = "white")
  ) +
  ggplot2::labs(x = "Assists/60", y = "Goals/60",
       title = "2000-01 Wings v Avs, regular season stats",
       subtitle = "min. 300 minutes",
       caption = "data pulled from hockey-reference.com using hockeyR")
```
