---
title: "Scraping hockey-reference.com"
author: "Daniel Morse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hockey-ref-scrapers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/vignettes-",
  out.width = "75%",
  dpi = 750,
  fig.width = 7,
  fig.height = 4.67
)
```

```{r setup}
library(hockeyR)
`%>%` <- magrittr::`%>%`
```

###   Hockey-Reference scrapers

These functions scrape data from [hockey-reference.com](https://www.hockey-reference.com).

Grab every team's win-loss record in any season going back to 1918 with the `get_team_records()` function

```{r records}
get_team_records(1967) %>%
  dplyr::arrange(-w) %>%
  dplyr::select(team_name, team_abbr, season, overall, w, l, otl, st_points)
```

You can also get stats down to the player-level with `get_player_stats_hr()`. This function defaults to the player's career statistics, but you can enter a specific season or range of seasons as well. Note that the season references the year the specific season ended (ie the 2021-22 season should be entered as 2022)

```{r player stats}
get_player_stats_hr(player_name = "Wayne Gretzky", season = 1982) %>%
  dplyr::select(player, age, season_full, tm, gp, g, a, pts)
```

Ever wonder who the most prolific goal-scorer was to wear a specific number? Use `get_jersey_players()` in conjunction with `get_player_stats_hr()` to find out:

```{r jerseys}
# get every player to wear the desired number
df <- get_jersey_players(98)

# get their statistics from the year they wore that sweater
df2 <- purrr::map2_dfr(
  .x = df$player,
  .y = df$season,
  ~get_player_stats_hr(player_name = .x, season = .y)
  )

# who had the most goals?
dplyr::arrange(df2, dplyr::desc(g)) %>%
  dplyr::select(player, tm, season_full, gp, g, a, pts)
```

You can use the data to make plots with actual team colors and logos as well using the `team_logos_colors` file included with the package.

```{r plot-example}
# add colors & logos
df3 <- df2 %>%
  dplyr::group_by(player, season_full) %>%
  # this part is just to get both of Mete's teams into one row
  dplyr::summarize(
    gp = sum(gp),
    pts = sum(pts),
    pts_gm = pts/gp,
    tm = utils::tail(tm, n=1),
    .groups = "drop"
  ) %>%
  dplyr::mutate(player_season = glue::glue("{player}\n{season_full}")) %>%
  dplyr::left_join(team_logos_colors, by = c("tm" = "team_abbr"))

# make a bar chart
df3 %>%
  ggplot2::ggplot(ggplot2::aes(stats::reorder(player_season, -pts_gm), pts_gm)) +
  ggplot2::geom_col(fill = df3$team_color1, color = df3$team_color2) +
  ggimage::geom_image(
    ggplot2::aes(y = pts_gm + .027, image = team_logo_espn),
    size = .07, asp = 1.5
  ) +
  ggplot2::geom_text(ggplot2::aes(y = 0.01, label = player_season),
            color = "white", angle = 90, hjust = 0) +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::theme(
    panel.background = ggplot2::element_rect(fill = "black"),
    plot.background = ggplot2::element_rect(fill = "black"),
    panel.grid.major.x = ggplot2::element_blank(),
    axis.text.x = ggplot2::element_blank(),
    axis.ticks.x = ggplot2::element_blank(),
    axis.text.y = ggplot2::element_text(color = "white"),
    title = ggplot2::element_text(color = "white")
  ) +
  ggplot2::labs(x = NULL, y = "Points per game",
       title = "Most productive NHL seasons wearing #98",
       caption = "data pulled from hockey-reference.com using hockeyR")
```
The most reliable way to grab player stats is through the `get_rosters()` function. The `get_player_stats_hr()` function still has some kinks to be worked out, as it works based on the player name (and messes up if you want to differentiate the Sebastian Ahos).

With `get_rosters()`, you can look up the current roster for any team in the league or the roster at season's end for any prior season. By default, it will only pull basic player info (name, age, height & weight, etc), but you can grab all the basic counting stats by setting `include_stats` to `TRUE`.

```{r rosters}
player_stats <- get_rosters(c("COL","Detroit red wings"), season = 2001, include_stats = TRUE) %>%
  dplyr::mutate(
    g_60 = 60 * g / toi,
    a_60 = 60 * a /toi,
    p_60 = 60 * pts / toi
  ) %>%
  dplyr::filter(toi >= 300) %>%
  dplyr::left_join(team_logos_colors, by = "team_abbr")

top_performers <- dplyr::filter(
      player_stats,
      p_60 >= dplyr::arrange(player_stats, -p_60) %>%
        dplyr::slice(10) %>%
        dplyr::pull(p_60)
      )

player_stats %>%
  ggplot2::ggplot(ggplot2::aes(a_60,g_60)) +
  ggplot2::geom_hline(yintercept = 60 * sum(player_stats$g) / sum(player_stats$toi),
             linetype = "dashed", color = "black") +
  ggplot2::geom_vline(xintercept = 60 * sum(player_stats$a) / sum(player_stats$toi),
             linetype = "dashed", color = "black") +
  #geom_point(aes(size = toi), show.legend = FALSE,
  #           color = player_stats$team_color_alt1, alpha = .8) +
  ggimage::geom_image(ggplot2::aes(image = team_logo_espn),
                      size = 0.07, asp = 1.5) +
  ggrepel::geom_text_repel(
    data = top_performers,
    ggplot2::aes(label = player),
    color = top_performers$team_color_alt1
  ) +
  ggplot2::scale_y_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::scale_x_continuous(breaks = scales::pretty_breaks()) +
  ggplot2::theme(
    panel.background = ggplot2::element_rect(fill = "#708090"),
    plot.background = ggplot2::element_rect(fill = "#708090"),
    title = ggplot2::element_text(color = "white")
  ) +
  ggplot2::labs(x = "Assists/60", y = "Goals/60",
       title = "2000-01 Wings v Avs, regular season stats",
       subtitle = "min. 300 minutes",
       caption = "data pulled from hockey-reference.com using hockeyR")
```
